# Data Modification Work space

## \> Import Data

```{r}
crop <- read.csv("../data/Crop_recommendation.csv")
print(head(crop))
```

```{r}
str(crop)
```

### Crops currently In the table

```{r}
cropsLabel <- unique(crop$label)
print(cropsLabel)
```

# 

## \> Load Kenyan Cash Crops To be Used

| Original Crop | Matched Kenyan Export Crop | Conditions Rationale                                                         |
|---------------|---------------|------------------------------------------|
| Rice          | Tea                        | High rainfall, humid, acidic soil (pH \~5.5-6.5)                             |
| Maize         | Coffee                     | Warm climate, moderate rainfall, slightly acidic to neutral soil             |
| Chickpea      | Avocado                    | Warm to hot climates, moderate rainfall, neutral to slightly acidic soil     |
| Kidneybeans   | Macadamia Nuts             | Warm climates, moderate to high rainfall, slightly acidic soil               |
| Pigeonpeas    | French Beans               | Warm climates, moderate rainfall, similar legume characteristics             |
| Mothbeans     | Snow Peas                  | Grows well in cool climates with moderate rainfall                           |
| Mungbean      | Passion Fruit              | Warm, humid, tropical environments, moderate to heavy rainfall               |
| Blackgram     | Mango                      | Warm, tropical climate, moderate rainfall, slightly acidic soil              |
| Lentil        | Pineapple                  | Tropical climates with good rainfall, grows well in acidic soils             |
| Pomegranate   | Flowers (Roses)            | Moderate climates, well-drained soil, moderate rainfall                      |
| Banana        | Cabbage                    | Grows in warm climates with moderate to heavy rainfall                       |
| Mango         | Sugarcane                  | Tropical and subtropical climates, needs high rainfall and warm temperatures |
| Grapes        | Cashew Nuts                | Warm, dry conditions with moderate rainfall, grows in well-drained soil      |
| Watermelon    | Tomatoes                   | Warm climates, moderate water needs, similar soil preferences                |
| Muskmelon     | Spinach                    | Prefers warm temperatures and moderate moisture                              |
| Apple         | Carrots                    | Grows in cooler climates with moderate rainfall                              |
| Orange        | Coconuts                   | Thrives in tropical climates with heavy rainfall and warm temperatures       |
| Papaya        | Sisal                      | Grows in semi-arid to dry conditions, requires warm temperatures             |
| Coconut       | Sesame Seeds               | Thrives in warm, dry conditions with moderate rainfall                       |
| Cotton        | Tobacco                    | Grows in warm to hot climates with moderate rainfall                         |
| Jute          | Chillies                   | Grows in warm climates with moderate humidity and rainfall                   |
| Coffee        | Pyrethrum                  | Grows well in high altitudes with cooler temperatures                        |

```{r}
# Mapped Kenyan export crops based on environmental conditions

kenya_export_crops <- c("Tea","Coffee","Avocado","Macadamia Nuts","French Beans","Snow Peas","Passion Fruit","Mango","Pineapple","Flowers (Roses)","Cabbage","Sugarcane","Cashew Nuts","Tomatoes","Spinach","Carrots","Coconuts","Sisal","Sesame Seeds","Tobacco","Chillies","Pyrethrum")

```

```{r}
print(kenya_export_crops)
```

```{r}

# Ensure there are no duplicates in the Kenyan exports list
if (length(unique(kenya_export_crops)) == length(kenya_export_crops)) {
    print("No duplicates in the Kenyan export crops.")
} else {
    print("There are duplicates in the Kenyan export crops!")
}

```

```{r}
# Print original and mapped crop labels to verify
for (i in 1:length(cropsLabel)) {
  cat("Original Crop:", cropsLabel[i], "->", "Kenyan Export Crop:", kenya_export_crops[i], "\n")
}
```

```{r}
# Create a named vector for easy replacement (map original crops to Kenyan crops)
crop_mapping <- setNames(kenya_export_crops, cropsLabel)


# Replace the 'label' column in the data frame with the mapped values
crop$label <- crop_mapping[crop$label]

# Print the first few rows to verify the replacements
print(head(crop))
```

# 

## \> Export The new modified Kenyan Cash crops Data

```{r}
# Specify the file name
output_file <- "../data/kenyan_cash_crops_conditions.csv"

# Export the data to a CSV file
write.csv(crop, file = output_file, row.names = FALSE)

# Confirmation message
cat("Data has been successfully exported to", output_file, "\n")
```

# 

# 

# Data Cleaning Work space

### Load Data

```{r}
kenyancrops <- read.csv("../data/kenyan_cash_crops_conditions.csv")

```

### Check for Missing values

```{r}
missing_vals <- sapply(kenyancrops, function(x) sum(is.na(x)))
print(missing_vals)
```

### Check for Duplicate Values

```{r}
duplicate_rows <- kenyancrops[duplicated(kenyancrops), ]

# Count the number of duplicate rows
num_duplicates <- nrow(duplicate_rows)

# Print the number of duplicate rows and the duplicate rows themselves
print(num_duplicates)
print(duplicate_rows)
```

### Count Unique Values in a Column

```{r}
label_counts <- table(kenyancrops$label)

print(label_counts)
```

# 

# 

# Data Visualization Work space

```{r}
# Packages Installation and Loading

#install.packages("ggplot2")

library(ggplot2)
```

### Bar Plot for Crop Counts

```{r}
#label_counts <- table(kenyancrops$label)  # Count unique values in 'label'

ggplot(data = as.data.frame(label_counts), aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Counts of Unique Crops", x = "Crop", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Scatter Plot for Rainfall vs. Humidity

```{r}

ggplot(kenyancrops, aes(x = rainfall, y = humidity, color = label)) +
  geom_point() +
  labs(title = "Scatter Plot of Rainfall vs Humidity", x = "Rainfall", y = "Humidity") +
  theme_minimal()

```

### Box Plot for Temperature by Crop

```{r}
# Create a box plot for temperature by crop
ggplot(kenyancrops, aes(x = label, y = temperature, fill = label)) +
  geom_boxplot() +
  labs(title = "Box Plot of Temperature by Crop", x = "Crop", y = "Temperature") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Histogram for Soil pH

```{r}
# Create a histogram for soil pH
ggplot(kenyancrops, aes(x = ph)) +
  geom_histogram(binwidth = 0.5, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Soil pH", x = "Soil pH", y = "Count") +
  theme_minimal()

```

# 

# 

# 

# MODEL WORKSPACE

### Load necessary libraries

```{r}
#install.packages(c("dplyr", "caret", "randomForest", "e1071"))

library(dplyr)
#library(caret)
#library(randomForest)
library(e1071)
```

> ## 1. Encoding the crops

```{r}
# Encoding the Kenyan crops
crop_dict <- list(
    'Tea' = 1, 'Coffee' = 2, 'Avocado' = 3, 'Macadamia Nuts' = 4, 
    'French Beans' = 5, 'Snow Peas' = 6, 'Passion Fruit' = 7, 
    'Mango' = 8, 'Pineapple' = 9, 'Flowers (Roses)' = 10, 'Cabbage' = 11, 
    'Sugarcane' = 12, 'Cashew Nuts' = 13, 'Tomatoes' = 14, 'Spinach' = 15, 
    'Carrots' = 16, 'Coconuts' = 17, 'Sisal' = 18, 'Sesame Seeds' = 19, 
    'Tobacco' = 20, 'Chillies' = 21, 'Pyrethrum' = 22
)

# Apply encoding
kenyancrops$crop_num <- unlist(crop_dict[kenyancrops$label])

# Drop the 'label' column
kenyancrops <- kenyancrops %>% select(-label)
```

## 2. Train Test Split

```{r}
set.seed(42)
# Split the data into training and testing sets (80% training, 20% testing)
trainIndex <- createDataPartition(kenyancrops$crop_num, p = 0.8, list = FALSE)
trainData <- kenyancrops[trainIndex,]
testData <- kenyancrops[-trainIndex,]

# Separate features (X) and target (Y)
x_train <- trainData %>% select(-crop_num)
y_train <- trainData$crop_num
x_test <- testData %>% select(-crop_num)
y_test <- testData$crop_num

```

## 3. Feature Scaling

```{r}
# Scale the features using Min-Max Scaling
preProcess_scale <- preProcess(x_train, method = 'range')
x_train_scaled <- predict(preProcess_scale, x_train)
x_test_scaled <- predict(preProcess_scale, x_test)

```

## 4. Model Training

```{r}
# Initialize models
models <- list(
    lg = train(x_train_scaled, y_train, method = "glm"),
    gnb = train(x_train_scaled, y_train, method = "naive_bayes"),
    svc = train(x_train_scaled, y_train, method = "svmRadial"),
    knn = train(x_train_scaled, y_train, method = "knn"),
    dt = train(x_train_scaled, y_train, method = "rpart"),
    rf = train(x_train_scaled, y_train, method = "rf"),
    adaboost = train(x_train_scaled, y_train, method = "adaboost")
)

# Evaluate the models
for(name in names(models)) {
    model <- models[[name]]
    y_pred <- predict(model, x_test_scaled)
    accuracy <- confusionMatrix(y_pred, y_test)$overall['Accuracy']
    print(paste(name, "accuracy:", accuracy))
}

```

## 5. Predictive System Testing

```{r}
# Fit the RandomForest model for predictions
rfc <- randomForest(x_train_scaled, y_train)
y_pred <- predict(rfc, x_test_scaled)
accuracy <- confusionMatrix(y_pred, y_test)$overall['Accuracy']
print(paste("Random Forest accuracy:", accuracy))

# Define a predictive function
recommendation <- function(N, P, k, temperature, humidity, ph, rainfall) {
    features <- data.frame(N = N, P = P, k = k, temperature = temperature,
                           humidity = humidity, ph = ph, rainfall = rainfall)
    features_scaled <- predict(preProcess_scale, features)
    prediction <- predict(rfc, features_scaled)
    return(prediction)
}

# Test the predictive system
N <- 20
P <- 30
k <- 40
temperature <- 40
humidity <- 20
ph <- 30
rainfall <- 50

predict_crop <- recommendation(N, P, k, temperature, humidity, ph, rainfall)

# Crop dictionary for output
crop_dict_rev <- c(1 = 'Tea', 2 = 'Coffee', 3 = 'Avocado', 4 = 'Macadamia Nuts', 5 = 'French Beans',6 = 'Snow Peas', 7 = 'Passion Fruit', 8 = 'Mango', 9 = 'Pineapple', 10 = 'Flowers (Roses)', 11 = 'Cabbage', 12 = 'Sugarcane', 13 = 'Cashew Nuts', 14 = 'Tomatoes', 15 = 'Spinach', 16 = 'Carrots', 17 = 'Coconuts', 18 = 'Sisal', 19 = 'Sesame Seeds', 20 = 'Tobacco', 21 = 'Chillies', 22 = 'Pyrethrum')

if (predict_crop %in% names(crop_dict_rev)) {
    print(paste(crop_dict_rev[as.character(predict_crop)], "is the best crop to be cultivated."))
} else {
    print("Sorry, we are unable to recommend a crop for this environment.")
}

```

# 

# 

# 

# Save the `rf` model along with the pre processing scaler for use in the Shiny app

```{r}
# Assuming 'rf' is the best-performing model based on accuracy
best_model <- models$rf

# Save the trained RandomForest model to an RDS file
saveRDS(best_model, file = "../models/random_forest_model.rds")

# Save the pre-processing scaler
saveRDS(preProcess_scale, file = "../models/preprocess_scaler.rds")

```
